Assessing Infant Mortality Rate: Problems stemming from Household Living Conditions, Women's Education and Health

### 2022-04-27 outputs/paper/paper.Rmd
final edits for clarity

---
title: "Assessing Infant Mortality Rate: Problems stemming from Household Living Conditions, Women's Education and Health"
subtitle: "Trends in India from the 1998 Demographic Health Survey"
author: 
  - Jacob Yoke Hong Si
thanks: "Code and data are available at: https://github.com/jacobyokehongsi/Assessing_Infant_Mortality_Rate_in_1998_India"
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "What areas can be improved in order to promote the well-being of women in India and hence, reduce the infant mortality rate? Utilizing the data from the 1998-1999 India National Family Health Survey provided by the Demographic and Health Survey (DHS) program, we look to depict the demographics of Indian women and infants in different states of India. We have found that the root causes of poor infant mortality rates stem from having poor living conditions that affect the likelihood of women to attain education and understand the importance of antenatal care and birth delivery assistance. We also explore other factors such as inheritable traits (unhealthy body weight and anaemia disease) as well as an infant's diet. These factors are crucial in the development of an infant and the reduction of the infant mortality rate. \\par \\textbf{Keywords:} demographics, health, living conditions, infant mortality rate, education, antenatal care, body weight, anaemia, diet, birth delivery assistance."
output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references.bib
---

@@

I use R [@citeR] and R packages tidyverse [@tidy] and pdftools [@pdftools] to extract, preprocess and analyze the data. When preprocessing the data, I also use R package pointblank [@pointblank] to put together tests for class and content to ensure that the dataset passes. The key features analyzed in this paper are shown in table 1 and 2, which is made using R package knitr [@knitr].

@@

# Results
The figures are made using R [@citeR], R package tidyverse [@tidy] and Lucidchart [@lucidchart]. The variables explored provides numeric values that varies between the states of India.

@@

Figure \@ref(fig:attsch) depicts the distribution of females age 6-14 attending school (in percent) across the different states of India. Here, we observe that majority of females attending school, are from states including Kerala, Himachal Pradesh and Goa whereas majority of females not attending school are from Bihar, Rajasthan and Uttar Pradesh. In the states where the majority of females are not attending school, we can infer that these states tend to have a larger population, especially Uttar Pradesh.

@@

Figure \@ref(fig:watersch) depicts the relationship between households with drinking water piped or from hand pump and females age 6-14 attending school (in percent). Here, we observe that as the percent of households drinking water piped or from hand pump increases, the percent of females age 6-14 attending school decreases at a moderate level which is odd. Therefore, we further analyze this strange trend by investigating the relationship between population and households with drinking water. Looking at Figure \@ref(fig:waterpop), as the percent of households with drinking water piped or from hand pump increases, the population increases. Thus, this aligns with the fact that in Figure \@ref(fig:attsch), states with higher populations are also states where the majority of females are not attending school.

@@

Figure \@ref(fig:schhealth) depicts the relationship between females age 6-14 attending school and women involved in decisions about their own healthcare (in percent). Here, we observe that as the percent of females age 6-14 attending school increases, the percent of women involved in decisions about their own healthcare increases. This suggests that female individuals that are educated are more likely to take care of their own health and live a healthier lifestyle.

@@

## Mothers that receive antenatal check-up, antenatal care and birth delivery assistance
Here, we explore how mothers that have antenatal check-up affect the antenatal care (supplements and injections) they receive for births three years preceding the survey.

@@

Figures \@ref(fig:anteiron) and \@ref(fig:anteinject) depict the relationships between mothers receiving at least one antenatal check-up with mothers receiving iron and folic acid tablets or syrup as well as two or more tetanus toxoid injections (in percent) respectively. We observe that as the percent of mothers receiving at least one antenatal check-up increases, the percent of mothers receiving two or more tetanus toxoid injections as well as iron and folic acid tablets or syrup increases. This suggests that as mothers attend antenatal check-ups, doctors are likely to prescribe the mothers with tetanus toxoid injections as well as iron and folic acid supplements which boosts the health of the mother for a good birth delivery.

@@

Figures \@ref(fig:antemed) and \@ref(fig:antehealth) depict the relationships between mothers receiving at least one antenatal check-up with deliveries in medical institutions as well as deliveries assisted by a health professional (in percent) respectively. We observe that as the percent of mothers receiving at least one antenatal check-up increases, the percent of deliveries in medical institutions as well as deliveries assisted by a health professional increases. This suggests that when mothers go for their antenatal check-up, they likely have to approach a medical institution and have their delivery conducted within the institution and/or by a health professional.

@@

### Antenatal care and Delivery Methods
Here, we explore how antenatal care such as supplements and injections as well as delivery methods affect the infant mortality rate for births in the three years preceding the survey.

@@

Figures \@ref(fig:ironmortal) and \@ref(fig:injectmortal) depicts the relationship between mothers receiving iron and folic acid tablets or syrup, and two or more tetanus toxoid injections (in percent) with infant mortality rate respectively. Here, we observe that as the percent of mothers receiving iron and folic acid tablets or syrup and two or more tetanus toxoid injections increases, the infant mortality rate decreases. This suggests that mothers that pay more attention to their own health and follow the doctor's antenatal prescriptions will lead to a reduction in deaths of infants when they are born.

@@

Figures \@ref(fig:medmortal) and \@ref(fig:promortal) depicts the relationship between deliveries in medical institutions and deliveries assisted by a health professional (in percent) with the infant mortality rate respectively. Here, we observe that as the percent of deliveries in medical institutions and deliveries assisted by a health professional increases, the infant mortality rate decreases. This implies that mothers that receive professional care during birth deliveries will also have a smooth child birth.

@@

Figures \@ref(fig:breastmortal) and \@ref(fig:breastfoodmortal) portrays the relationship between infant age 0-3 months exclusively breastfed, and infant age 6-9 months receiving breast milk and solid/mushy food (in percent) with the infant mortality rate respectively. Here, we observe that the percent of infant age 0-3 months that are exclusively breastfed does not have a significant impact on infant mortality rate. On the other hand, as the percent of infant age 6-9 months receiving breast milk and solid/mushy food increases, the infant mortality rate decreases. This indicates that regardless of breast milk or formula milk fed to infants at age 0-3 months, infant mortality rate is not affected. However, as infant grow to the age of 6-9 months, they should be provided solid/mushy food on top of breast milk to support their growth and reduce the risk of mortality.

### Relationship between a women's body mass index (BMI), infant's weight and mortality rate

@@

Figure \@ref(fig:bmiunder) depicts the relationship between women with BMI below 18.5 kg/m and infant under age three years that are underweight (in percent). Women that have BMI below 18.5kg/m is considered to be underweight [@bmi]. Here, we observe that as the percent of women with BMI below 18.5 kg/m increases, the percent of infant under age three years that are underweight increases. This implies that mothers that are skinny and malnutritioned would also lead to offsprings that are underweight.

@@

Figure \@ref(fig:undermortal) depicts the relationship between infant under age three years that are underweight (in percent) and infant mortality rate. Here, we observe that as the percent of infant under age three that are underweight increases, the infant mortality rate increases. From this we can infer that infant who are not underweight tends to be in good health which reduces infant mortality rate.

### Relationship between women with anaemia, infant with anaemia and infant mortality rate

@@

Figure \@ref(fig:womenchildanaem) depicts the relationship between women age 15-49 with any anaemia and infant age 6-35 months with any anaemia (in percent). Here, we observe that as the percent of women age 15-49 with any anaemia increases, the percent of infant age 6-35 months with any anaemia increases. This is coherent with the fact that since anaemia such as hemolytic anaemia can be inherited, mothers could pass on the anaemia to their offspring.

@@

Figure \@ref(fig:childmortal) depicts the infant age 6–35 months with any anaemia (in percent) and infant mortality rate. Here, we observe that as the percent of infant age 6–35 months with any anaemia increases, the infant mortality rate increases. From this we can infer that infant who have the anaemia disease would have symptoms that lead to critical health conditions and death.

# Discussion
The ultimate goal of the paper is to assist policy makers in implementing appropriate and efficient socio-economic policies to help promote the steady growth of infants and reduce infant mortality rate in India. We will first discuss the root of the issue which is the household living environment, women's education and their healthcare decisions which lead them to attend antenatal check-ups. Then, we examine the numerous factors that are affected by conducting antenatal check-ups which include antenatal supplements and injections, as well as birth delivery assistance. These are some of the medical factors that directly impact the infant mortality rate. Additionally, we also look at the importance of infant's meals for their growth. Lastly, we examine inheritable traits such as anaemia and unhealthy body weight that could be passed on from mother to infant and how they impact the infant mortality rate. The full data can be found in Appendix B.

## Household Living Environment, Women's Education and their Healthcare Decisions 
The household living environment of individuals including access to electricity, clean water and sanitation facilities is crucial in providing a comfortable and productive living space. As observed previously, households with better living conditions will also lead to women who are able to attend school and be well educated. In our figures, we could infer that households in India with poorer living conditions also have fewer women attending school. Thus, emphasis should be put in developing the living conditions within households in India, especially states with poorer living conditions such as Bihar. In Bihar, only 18.2% of the households have access to electricity and 54.1% of females attending school. With such poor basic needs and living standards, there is no doubt that the women is unable to attend school. By developing the living conditions, it will in turn improve the well-being of the population and encourage women to attend school. Another proposal could also be to provide direct subsidies for students in hopes to attain a well educated population.

@@

## Antenatal check-ups, antenatal care and delivery methods to promote mothers and infants health
Attending antenatal check-ups is key for protecting the health of women and their unborn infants. During antenatal check-ups, it is likely that doctors prescribe supplements and procedures needed in order for a smooth child birth. This includes iron and folic acid supplements that supports the development of the placenta and fetus as well as preventing neural tube defects respectively [@ironfolic]. In addition, tetanus toxoid injections help prevent neonatal tetanus [@tetanus]. Furthermore, partaking in check-ups will also increase the likelihood of having child deliveries in medical institutions and/or by a health professional since the check-ups are conducted in a similar fashion. Child deliveries in this manner are safe and sanitary, ensuring that complications are minimized when the baby is delivered. In our data, we observe that in states where mothers often receive at least one antenatal check-up such as Goa (99.0% of mothers receiving at least one antenatal check-up), they also pay more attention to antenatal care where mothers receive two or more tetanus toxoid injections (86.1%), iron and folic acid tablets or syrups (94.7%) as well as having their infant delivered in medical institutions (90.8%) and/or by a health professional (90.8%).

## Meals of infants and their effects on infant mortality rate
The diet of an infant is one of the most important fundamentals in a child's health. An infant has to be well-fed in order to receive the nutrients and vitamins required for a healthy body. To recap, we observe that for infant age 0-3 months, being exclusively breastfed does not have a significant impact on the infant mortality rate. On the other hand, for infant age 6-9 months, it is important that they receive solid/mushy food to supplement the breast milk they are consuming. States such as Rajasthan with low percent of infant age 6-9 months receiving breast milk and solid/mushy food (17.5%) also have high infant mortality rates (80.4%). As a child grows to more than 6 months old, the nutrients supplied by the breast milk is simply insufficient to develop a healthy immune system. Thus, it is vital that infants are supplied with ample amount of nutritious solid foods on top of breast/formula milk to promote a healthy growth.

## Unhealthy body weight and anaemia of mothers affecting the infant mortality rate
Lastly, we will explore how inheritable traits that mothers possess such as unhealthy body weight and diseases (anaemia) are inherited by their infant and would be detrimental to their health. In the state Orissa, the percent of women who are underweight is roughly 48.0% which corresponds to 54.4% of children being underweight. Hence, women that are underweight would increase the likelihood of their offsprings being underweight and malnourished.

On the other hand, infants could also inherit anaemia diseases from their mother. Anaemia is a condition in which the body does not have sufficient healthy red blood cells to carry adequate oxygen around the body. Symptoms of anaemia include tiredness, shortness of breath, irregular heartbeats and more [@anaemia]. As infants are born with this disease, they are fatigued and sick leading to their poor health and death. In the northeast state of Assam, the percent of women age 15-49 with any anaemia is approximately 69.7% which also corresponds to the high anaemia rates of 63.2% in infants age 6-35 months. The percentage of infants with anaemia is relatively close to the percentage of women with anaemia. From this, we can infer that infants are most likely to have inherited this disease from their mother.

@@

There are a few weaknesses in the paper that I will evaluate. Firstly, the data is not collected by myself and hence, we only know as much about the methodology behind it as we've been provided. Despite this being a potentially major issue, the IIPS did more than their due diligence by providing supple information in explaining every aspects of the survey including the survey methodology and sampling techniques as well as documenting detailed variable names. However, there could be potential biases that is unaccounted for by the IIPS that includes the sampling or surveying process and ethical considerations that impacts the data collection in which we have no information about them.

Another possible weakness in our paper is the uncertainty in which the factors are individually correlated to the infant mortality rate. For instance, infant that do not receive solid/mushy food is not the sole reason for the deaths in the infant. Instead, it is the amalgamation of all the different factors analyzed that leads to the deaths of infants. As a next step for this issue, we could conduct statistical tests such as t-tests to determine if individual variables are statistically significant when regressed on the infant mortality rate. Apart from the aforementioned variables, there are factors such as the household income, family size etc. that could affect a female's ability of attending school. Hence, by collecting more data on more variables, we would be able to make even more concrete conclusions with regards to the factors that contribute to infant mortality.

To summarize, there are numerous factors that could affect infant mortality rate which includes the weight of a child, whether a child has anaemia, diet of the child and the prenatal health of the mother. The health of the mother is affected by her education background and the living conditions she grew up in. Hence, it is crucial for the government to develop policies that solves the problem at its root by first improving the living standards of households, starting with states such as Bihar with the poorest living conditions. The government could also build more schools as well as subsidize education expenses for poor families which encourages them to send their children to school. Having good living conditions and educated women would improve the health of the mother and her offspring, resulting in a healthier future generation of India.

>>>>

---
title: "Assessing Infant Mortality Rate: Problems stemming from Household Living Conditions, Women's Education and Health"
subtitle: "Trends in India from the 1998 Demographic Health Survey"
author: 
  - Jacob Yoke Hong Si
thanks: "Code and data are available at: https://github.com/jacobyokehongsi/Assessing_Infant_Mortality_Rate_in_1998_India"
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "What areas can be improved in order to promote the well-being of women in India and hence, reduce the infant mortality rate? Utilizing the data from the 1998-1999 India National Family Health Survey provided by the Demographic and Health Survey (DHS) program, we look to depict the demographics of Indian women and infants in different states of India. We have found that the root causes of poor infant mortality rates stem from having poor living conditions that affect the likelihood of women to attain education and understand the importance of antenatal care and birth delivery assistance. We also explore other factors such as potentially inheritable traits (unhealthy body weight and anaemia disease) as well as an infant's diet. These factors are crucial in the development of an infant and the reduction of the infant mortality rate. \\par \\textbf{Keywords:} demographics, health, living conditions, infant mortality rate, education, antenatal care, body weight, anaemia, diet, birth delivery assistance."
output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references.bib
---

@@

We use R [@citeR] and R packages tidyverse [@tidy] and pdftools [@pdftools] to extract, preprocess and analyze the data. When preprocessing the data, We also use R package pointblank [@pointblank] to put together tests for class and content to ensure that the dataset passes. The key features analyzed in this paper are shown in table 1 and 2, which are made using R package knitr [@knitr].

@@

# Results
The figures are made using R [@citeR], R package tidyverse [@tidy] and Lucidchart [@lucidchart]. The variables explored provide numeric values that vary between the states of India.

@@

Figure \@ref(fig:attsch) depicts the distribution of females age 6-14 attending school (in percent) across the different states of India. Here, we observe that majority of females attending school, are from states including Kerala, Himachal Pradesh and Goa whereas majority of females not attending school are from Bihar, Rajasthan and Uttar Pradesh. In the states where the majority of females are not attending school, we can also infer that these states tend to have a larger population, especially Uttar Pradesh.

@@

Figure \@ref(fig:watersch) depicts the relationship between households with drinking water piped or from hand pump and females age 6-14 attending school (in percent). Here, we observe that as the percent of households drinking water piped or from hand pump increases, the percent of females age 6-14 attending school decreases slightly which is odd. Therefore, we further analyze this strange trend by investigating the relationship between population and households with drinking water. Looking at Figure \@ref(fig:waterpop), as the percent of households with drinking water piped or from hand pump increases, the population increases. Thus, this aligns with the fact that in Figure \@ref(fig:attsch), states with higher populations are also states where the majority of females are not attending school.

@@

Figure \@ref(fig:schhealth) depicts the relationship between females age 6-14 attending school and women involved in decisions about their own healthcare (in percent). Here, we observe that as the percent of females age 6-14 attending school increases, the percent of women involved in decisions about their own healthcare increases. This suggests that female individuals who are educated are more likely to take care of their own health and live a healthier lifestyle.

@@

## Mothers that receive antenatal check-up, antenatal care and birth delivery assistance
Here, we explore how mothers that have antenatal check-up affect the antenatal care (supplements and injections) and birth delivery assistance they receive for births three years preceding the survey.

@@

Figures \@ref(fig:anteiron) and \@ref(fig:anteinject) depict the relationships between mothers receiving at least one antenatal check-up with mothers receiving iron and folic acid tablets or syrup as well as two or more tetanus toxoid injections (in percent) respectively. We observe that as the percent of mothers receiving at least one antenatal check-up increases, the percent of mothers receiving iron and folic acid tablets or syrup as well as two or more tetanus toxoid injections increases. This suggests that as mothers attend antenatal check-ups, doctors are likely to prescribe the mothers with iron and folic acid supplements as well as tetanus toxoid injections which boosts the health of the mother and her newborn infant.

@@

Figures \@ref(fig:antemed) and \@ref(fig:antehealth) depict the relationships between mothers receiving at least one antenatal check-up with deliveries in medical institutions as well as deliveries assisted by a health professional (in percent) respectively. We observe that as the percent of mothers receiving at least one antenatal check-up increases, the percent of deliveries in medical institutions as well as deliveries assisted by a health professional increases. This suggests that when mothers go for their antenatal check-up, they likely have to approach a medical institution and have their deliveries conducted within the institution and/or by a health professional.

@@

### Antenatal care and Delivery Assistance
Here, we explore how antenatal care such as supplements and injections as well as delivery assistance in the three years preceding the survey affect the infant mortality rate for births.

@@

Figures \@ref(fig:ironmortal) and \@ref(fig:injectmortal) depicts the relationship between mothers receiving iron and folic acid tablets or syrup, and two or more tetanus toxoid injections (in percent) with infant mortality rate respectively. Here, we observe that as the percent of mothers receiving iron and folic acid tablets or syrup and two or more tetanus toxoid injections increases, the infant mortality rate decreases. This suggests that mothers who pay more attention to their own health and follow the doctor's antenatal prescriptions will lead to a reduction in deaths of infants after they are born.

@@

Figures \@ref(fig:medmortal) and \@ref(fig:promortal) depicts the relationship between deliveries in medical institutions and deliveries assisted by a health professional (in percent) with the infant mortality rate respectively. Here, we observe that as the percent of deliveries in medical institutions and deliveries assisted by a health professional increases, the infant mortality rate decreases. This implies that mothers who receive professional medical services during birth deliveries will also have a smooth child birth.

@@

Figures \@ref(fig:breastmortal) and \@ref(fig:breastfoodmortal) portrays the relationship between infant age 0-3 months exclusively breastfed, as well as infant age 6-9 months receiving breast milk and solid/mushy food (in percent) with the infant mortality rate respectively. Here, we observe that the percent of infant age 0-3 months that are exclusively breastfed does not have a significant impact on infant mortality rate. On the other hand, as the percent of infant age 6-9 months receiving breast milk and solid/mushy food increases, the infant mortality rate decreases. This indicates that regardless of breast milk or formula milk fed to infants at age 0-3 months, infant mortality rate is not affected. However, as infants grow to the age of 6-9 months, they should be provided solid/mushy food on top of breast milk to support their growth and reduce the risk of mortality.

### Relationship between a women's body mass index (BMI), infants' weight and mortality rate

@@

Figure \@ref(fig:bmiunder) depicts the relationship between women with BMI below 18.5 kg/m and infants under age three years that are underweight (in percent). Women that have BMI below 18.5kg/m are considered to be underweight [@bmi]. Here, we observe that as the percent of women with BMI below 18.5 kg/m increases, the percent of infants under age three years that are underweight increases. This implies that mothers that are skinny and malnutritioned would also lead to offsprings that are underweight.

@@

Figure \@ref(fig:undermortal) depicts the relationship between infants under age three years that are underweight (in percent) and infant mortality rate. Here, we observe that as the percent of infants under age three that are underweight increases, the infant mortality rate increases. From this we can infer that infants who are not underweight tend to be in good health which reduces infant mortality rate.

### Relationship between women with anaemia, infants with anaemia and infant mortality rate

@@

Figure \@ref(fig:womenchildanaem) depicts the relationship between women age 15-49 with any anaemia and infants age 6-35 months with any anaemia (in percent). Here, we observe that as the percent of women age 15-49 with any anaemia increases, the percent of infants age 6-35 months with any anaemia increases. This is coherent with the fact that since anaemia such as hemolytic anaemia can be inherited, mothers could pass on the anaemia to their offsprings.

@@

Figure \@ref(fig:childmortal) depicts the infants age 6–35 months with any anaemia (in percent) and infant mortality rate. Here, we observe that as the percent of infants age 6–35 months with any anaemia increases, the infant mortality rate increases. From this we can infer that infants who have the anaemia disease would have symptoms that lead to critical health conditions and death.

# Discussion
The ultimate goal of the paper is to assist policy makers in implementing appropriate and efficient socio-economic policies to help promote the steady growth of infants and reduce infant mortality rate in India. We will first discuss the root causes which are the household living environment, women's education and their healthcare decisions which lead them to attend antenatal check-ups. Then, we examine the numerous factors that are affected by conducting antenatal check-ups which include antenatal supplements and injections, as well as birth delivery assistance. These are some of the medical factors that directly impact the infant mortality rate. Additionally, we also look at the importance of infants' meals for their growth. Lastly, we examine potentially inheritable traits such as anaemia and unhealthy body weight that could be passed on from mother to infant and how they impact the infant mortality rate. The full data can be found in Appendix B.

## Household Living Environment, Women's Education and their Healthcare Decisions 
The household living environment of individuals including access to electricity, clean water and sanitation facilities is crucial in providing a comfortable and productive living space. As observed previously, households with better living conditions will also lead to women who are able to attend school and be well educated. In our figures, we could infer that households in India with poorer living conditions also have fewer women attending school. Thus, emphasis should be put in developing the living conditions within households in India, especially states with poorer living conditions such as Bihar. In Bihar, only 18.2% of the households have access to electricity and 54.1% of females attending school. With such poor basic necessities and living standards, there is no doubt that the women are unable to attend school. By developing the living conditions, it will in turn improve the well-being of the population and encourage women to attend school. Another proposal could also be to provide direct subsidies for needy students in hopes to attain a well educated population.

@@

## Antenatal check-ups, antenatal care and delivery assistance to promote mothers and infants health
Attending antenatal check-ups is key for protecting the health of women and their unborn infants. During antenatal check-ups, it is likely that doctors prescribe supplements and procedures needed in order for a smooth child birth. This includes iron and folic acid supplements that supports the development of the placenta and fetus as well as preventing neural tube defects respectively [@ironfolic]. In addition, tetanus toxoid injections help prevent neonatal tetanus [@tetanus]. Furthermore, partaking in check-ups will also increase the likelihood of having child deliveries in medical institutions and/or by a health professional since the check-ups are conducted in a similar fashion. Child deliveries in this manner are safe and sanitary, ensuring that complications are minimized when the baby is delivered. In our data, we observe that in states where mothers often receive at least one antenatal check-up such as Goa (99.0% of mothers receiving at least one antenatal check-up), they also pay more attention to antenatal care where mothers receive two or more tetanus toxoid injections (86.1%), iron and folic acid tablets or syrups (94.7%) as well as having their infant delivered in medical institutions (90.8%) and/or by a health professional (90.8%).

## Meals of infants and their effects on infant mortality rate
The diet of an infant is one of the most important fundamentals in a child's health. An infant has to be well-fed in order to receive the nutrients and vitamins required for a healthy body. To recap, we observe that for infants age 0-3 months, being exclusively breastfed does not have a significant impact on the infant mortality rate. On the other hand, for infants age 6-9 months, it is important that they receive solid/mushy food to supplement the breast milk they are consuming. States such as Rajasthan with low percent of infants age 6-9 months receiving breast milk and solid/mushy food (17.5%) also have a high infant mortality rate (80.4%). As an infant grows to more than 6 months old, the nutrients supplied by the breast milk are simply insufficient to develop a healthy immune system. Thus, it is vital that infants are supplied with ample amount of nutritious solid foods on top of breast/formula milk to promote a healthy growth.

## Unhealthy body weight and anaemia of mothers affecting the infant mortality rate
Lastly, we will explore how potentially inheritable traits that mothers possess, such as unhealthy body weight and anaemia disease, are inherited by their infants and would be detrimental to their health. In the state Orissa, the percent of women who are underweight is roughly 48.0% which corresponds to 54.4% of infants being underweight. Hence, women that are underweight would increase the likelihood of their offsprings being underweight and malnourished.

On the other hand, infants could also inherit anaemia diseases from their mother. Anaemia is a condition in which the body does not have sufficient healthy red blood cells to carry adequate oxygen around the body. Symptoms of anaemia include tiredness, shortness of breath, irregular heartbeats and more [@anaemia]. As infants are born with this disease, they are fatigued and sick leading to their poor health and death. In the northeast state of Assam, the percent of women age 15-49 with any anaemia is approximately 69.7% which also corresponds to the high anaemia rate of 63.2% in infants age 6-35 months. The percentage of infants with anaemia is relatively close to the percentage of women with anaemia. From this, we can infer that infants are most likely to have inherited this disease from their mother.

@@

There are a few weaknesses in the paper that I will evaluate. Firstly, the data was not collected by myself and hence, we only know as much about the methodology behind it as we've been provided. Despite this being a potentially major issue, the IIPS did more than their due diligence by providing supple information in explaining every aspects of the survey including the survey methodology and sampling techniques as well as documenting detailed variable names. However, there could be potential biases that were unaccounted for by the IIPS that included the sampling or surveying process, and ethical considerations that impacted the data collection in which we have no information about them.

Another possible weakness in our paper is the uncertainty in which the factors are individually correlated to the infant mortality rate. For instance, infants that do not receive solid/mushy food is not the sole reason for the deaths of infants. Instead, it is the amalgamation of all the different factors analyzed that leads to their deaths. As a next step for this issue, we could conduct statistical tests such as t-tests to determine if individual variables are statistically significant when regressed on the infant mortality rate. Apart from the aforementioned variables, there are factors such as the household income, number of schools in a state etc. that could affect a female's ability of attending school. Hence, by collecting more data on more variables, we would be able to make even more concrete conclusions with regards to the factors that contribute to infant mortality.

To summarize, there are numerous factors that could affect infant mortality rate which include the unhealthy weight of an infant, whether an infant has anaemia, diet of an infant and the prenatal health of a mother. The health of a mother is affected by her education background and the living conditions she grew up in. Hence, it is crucial for the government to develop policies that solves the fundamental problems by first improving the living standards of households, starting with states such as Bihar with the poorest living conditions. The government could also build more schools as well as subsidize education expenses for poor families which encourages them to send their children to school. Having good living conditions and educated women would improve the health of mothers and her offsprings, resulting in a healthier future generation of India.
